---
title: Benchmarking
subtitle: Ayush Noori
editor_options: 
  chunk_output_type: inline
---

```{r setup, include = FALSE}
# none
```

# Dependencies

Load requisite packages and define directories. Note that this script uses my personal utilities package `brainstorm`, which can be downloaded via `devtools::install_github("ayushnoori/brainstorm")`.

```{r load-packages, message=FALSE, warning=FALSE}
# data manipulation
library(data.table)
library(purrr)
library(magrittr)

# string manipulation
library(stringr)

# gene network inference
library(GENIE3)

# utility functions
library(brainstorm)
library(here)
```


Note that directories are relative to the R project path.

```{r define-directores}
# set directories
bdir = here("benchmarking")

# read DREAM5 data
ecoli_data = fread(here(bdir, "ecoli_data.csv"), header = TRUE)
ecoli_gene_names = fread(here(bdir, "ecoli_gene_names.csv"), header = TRUE)[["0"]]
ecoli_tf_names = fread(here(bdir, "ecoli_tf_names.csv"), header = TRUE)[["0"]]

# read test data split
test = fread(here(bdir, "test_dataset.csv"))
```

# GENIE3 Benchmarking

Use GENIE3 to construct connectivity graph. Prepare data in `exprMat` format. Running on multiple cores requires the `doParallel` and `doRNG` packages.

> "We chose the widely accepted random forest based GENIE3 along with information based ARACNE, CLR, and MRNET as the representative unsupervised methods in comparison."

```{r genie3-benchmarking}
# set row names
expr_mtx = as.data.frame(ecoli_data[, -1])
rownames(expr_mtx) = ecoli_data$Gene

# calculate regulators and targets
regulators = ecoli_tf_names
targets = ecoli_gene_names[!ecoli_gene_names %in% regulators]
```

Create weight matrix.

```{r weight-matrix}
# create weight matrix
weights = GENIE3(as.matrix(expr_mtx), regulators = regulators, targets = targets, verbose = T) # nCores = 10
```

Save to CSV.

```{r save-genie3}
# save to CSV
fwrite(as.data.frame(t(weights), row.names = colnames(weights)), here(bdir, "genie3_weight_matrix.csv"), row.names = TRUE)
```

# Information-Based Benchmarking

> "The implementation of ARACNE, CLR, and MRNET in R package `minet` is employed for analysis."

```{r information-benchmarking}
# load minet package
library(minet)

# transpose matrix
expr_mtx_t = t(expr_mtx)

# run MINET calculations
mr = minet(expr_mtx_t, method = "mrnet", estimator = "spearman")
ar = minet(expr_mtx_t, method = "aracne", estimator = "spearman")
clr = minet(expr_mtx_t, method = "clr", estimator = "spearman")

# save to CSV
# fwrite(mr, here(bdir, "mrnet_weight_matrix.csv"), row.names = TRUE)
# fwrite(ar, here(bdir, "aracne_weight_matrix.csv"), row.names = TRUE)
# fwrite(clr, here(bdir, "clr_weight_matrix.csv"), row.names = TRUE)
```

Check if matrices are symmetric.

```{r check-symmetric}
message("mrnet symmetric? ", all(mr == t(mr)))
message("aracne symmetric? ", all(ar == t(ar)))
message("clr symmetric? ", all(clr == t(clr)))
```

Filter matrices for targets as rows and regulators as columns.

```{r filter-matrices}
# calculate regulators and targets
regulators = ecoli_tf_names
targets = ecoli_gene_names[!ecoli_gene_names %in% regulators]

# filter matrices
mr_filter = mr[rownames(mr) %in% targets, colnames(mr) %in% regulators]
ar_filter = ar[rownames(ar) %in% targets, colnames(ar) %in% regulators]
clr_filter = clr[rownames(clr) %in% targets, colnames(clr) %in% regulators]

# write to output
fwrite(as.data.frame(mr_filter, row.names = rownames(mr_filter)), here(bdir, "mrnet_weight_matrix.csv"), row.names = TRUE)
fwrite(as.data.frame(ar_filter, row.names = rownames(ar_filter)), here(bdir, "aracne_weight_matrix.csv"), row.names = TRUE)
fwrite(as.data.frame(clr_filter, row.names = rownames(clr_filter)), here(bdir, "clr_weight_matrix.csv"), row.names = TRUE)
```